Draft-assinment1

### 2017-02-13 Draft-xuezhou-sjb.Rmd
added relationship

```{r setup, include=FALSE,warning=FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

@@

library(readr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(ggvis)

@@

data = read.csv("MERGED2012_13_PP.csv",header = T, na.strings = "NULL")

@@

variable = select(data, STABBR, PREDDEG, LOCALE2, MENONLY, WOMENONLY, CONTROL, ADM_RATE, SATMTMID, SATVRMID, SATWRMID, SAT_AVG,
                  NPT4_PUB,
                  NPT4_PRIV,
                  UG,
                  UG_NRA,
                  UG_UNKN,
                  UG_WHITENH,
                  UG_BLACKNH,
                  UG_API,
                  UG_AIANOLD,
                  UG_HISPOLD,
                  DEP_STAT_PCT_IND,
                  PAR_ED_PCT_1STGEN,
                  MD_FAMINC,
                  MEDIAN_HH_INC,
                  POVERTY_RATE,
                  PCTPELL,
                  PCTFLOAN,
                  GRAD_DEBT_MDN,
                  WDRAW_DEBT_MDN,
                  APPL_SCH_PCT_GE5,
                  C150_4,
                  MD_EARN_WNE_P6,
                  MN_EARN_WNE_P10,
                  RPY_3YR_RT_SUPP,
                  CDR3)

@@

```
From the output, we can tell that the highest value is 95.25% while the lowest value is 6.87%. From the mean value of 48.08%, we can tell that about for federally aided students in most colleges, around half (50.8%) of the students' parents did not complete the college. The standard deviation value of 12.2% indicated a relative high difference of this element among different colleges. Below is the code to generate a visual representation in the histgram. 
```{R}

  ggplot(data=data_clean , aes(PAR_ED_PCT_1STGEN)) + 
    geom_histogram(breaks=seq(0.1, 1, by =0.1), 
                   col="red", 
                   aes(fill=..count..),na.rm=TRUE)+
    labs(title="Histogram for PAR_ED_PCT_1STGEN") +
    labs(x="PAR_ED_PCT_1STGEN", y="Count") +
    theme(plot.title = element_text(hjust = 0.5))
```
From the histogram, we can tell the share of first generarion students is most frequently falling into the range between 0.5 and 0.6. The value with the value more than 0.8 are rare. The colleges with the share of first generarion students value less than 0.5 are much more than the ones with the value larger than 0.5. Thus the histogram is seriously left skewed.

Then we grouped by the data into the state level to take a further analysis.
```{R message=FALSE, warinig = FALSE}

data_clean %>%
  group_by(variable$STABBR ) %>%
    summarise(minPAR_ED_PCT_1STGEN=min(PAR_ED_PCT_1STGEN ,na.rm=TRUE),    
    maxPAR_ED_PCT_1STGEN=max(PAR_ED_PCT_1STGEN,na.rm=TRUE),medianPAR_ED_PCT_1STGEN=median(PAR_ED_PCT_1STGEN,na.rm=TRUE))  %>%
       arrange(medianPAR_ED_PCT_1STGEN)   %>%
{.}  ->PAR_ED_PCT_1STGEN_STATE

data_clean %>%
  group_by(variable$ctrF ) %>%
    summarise(minPAR_ED_PCT_1STGEN=min(PAR_ED_PCT_1STGEN ,na.rm=TRUE),    
    maxPAR_ED_PCT_1STGEN=max(PAR_ED_PCT_1STGEN,na.rm=TRUE),medianPAR_ED_PCT_1STGEN=median(PAR_ED_PCT_1STGEN,na.rm=TRUE))  %>%
       arrange(desc(medianPAR_ED_PCT_1STGEN))   %>%
{.}  ->PAR_ED_PCT_1STGEN_STATE

@@

Then we grouped by the data into the state level to take a further analysis.
```{R message=FALSE, warinig = FALSE}

data_clean %>%
  group_by(variable$STABBR) %>%
    summarise(minNPT4_COMBINE=min(NPT4_COMBINE ,na.rm=TRUE),    
    maxNPT4_COMBINE=max(NPT4_COMBINE,na.rm=TRUE),medianNPT4_COMBINE=median(NPT4_COMBINE,na.rm=TRUE))  %>%
       arrange(desc(medianNPT4_COMBINE))   %>%
{.}  ->NPT4_STATE

below is the code to generate stacked hisogram
``` {r}
data_clean %>%
  group_by(variable$STABBR, na.rm=TRUE) %>%
  ggvis(x=~NPT4_COMBINE,na.rm=TRUE) %>%
  layer_histograms(width=5, stack=TRUE, fill=~variable$STABBR,na.rm=TRUE)

data_clean %>%
  ggplot(aes(x=NPT4_COMBINE)) +
  geom_bar(binwidth=50,aes(fill=variable$ctrF,na.rm=TRUE))

data_clean %>%
  ggplot(aes(x=NPT4_COMBINE,na.rm=TRUE)) +
  geom_bar(binwidth=3000,aes(fill=variable$ctrF,na.rm=TRUE),breaks=seq(-20000, 90000, by =3000))

data_clean %>%
  ggplot(aes(x=NPT4_COMBINE,na.rm=TRUE)) +
  geom_bar(binwidth=50,aes(fill=variable$ctrF,na.rm=TRUE),breaks=seq(-20000, 90000, by =50))

@@

  #histogram for NPT_COM
  ggplot(data=data_clean , aes(DEP_STAT_PCT_IND)) + 
    geom_histogram(breaks=seq(0.1, 1, by =0.1), 
                   col="red", 
                   aes(fill=..count..),na.rm=TRUE)+
    labs(title="Histogram for DEP_STAT_PCT_IND") +
    labs(x="DEP_STAT_PCT_IND", y="Count") +
    theme(plot.title = element_text(hjust = 0.5))
```
From the histogram, we notice that the value between 0.7 to 0.8 are quite frequent. There was no obvious extrme value in the sample, thus the histogram is normally distributed.

@@

  ggplot(data=data_clean , aes(PCTPELL)) + 
    geom_histogram(breaks=seq(0, 1, by =0.1), 
                   col="red", 
                   aes(fill=..count..),na.rm=TRUE)+
    labs(title="Histogram for PCTPELL") +
    labs(x="PCTPELL", y="Count") +
    theme(plot.title = element_text(hjust = 0.5))
```
From the histogram, we are able to conclude that in most colleges, 40% to 50% of the undergraduate students are able to receive the Pell Grant. With little skewness, the histogram is resonalbe symmetric.

data_clean %>%
  group_by(variable$ctrF ) %>%
    summarise(minPCTPELL=min(PCTPELL ,na.rm=TRUE),    
    maxPCTPELL=max(PCTPELL,na.rm=TRUE),medianPCTPELL=median(PCTPELL,na.rm=TRUE))  %>%
       arrange(medianPCTPELL)   %>%
{.}  ->PCTPELL_CONTROL

@@

  ggplot(data=data_clean , aes(PCTFLOAN)) + 
    geom_histogram(breaks=seq(0, 1, by =0.1), 
                   col="red", 
                   aes(fill=..count..),na.rm=TRUE)+
    labs(title="Histogram for PCTFLOAN") +
    labs(x="PCTFLOAN", y="Count") +
    theme(plot.title = element_text(hjust = 0.5))
```
From the histogram, it's obvious that there are two peaks in the graph. The rates between 0- 10% and the rates between 60% to 80% are similarly frequent compated to the other ranges.

@@

data_clean %>%
  group_by(variable$ctrF ) %>%
    summarise(minPCTFLOAN=min(PCTFLOAN ,na.rm=TRUE),    
    maxPCTFLOAN=max(PCTFLOAN,na.rm=TRUE),medianPCTFLOAN=median(PCTFLOAN,na.rm=TRUE))  %>%
       arrange(medianPCTFLOAN)   %>%
{.}  ->PCTFLOAN_CONTROL

@@

```
From the result, the Private institutions have a much larger value than the public institutions. With a much higher costs burden, it makes sence that the low-income sutdents in private institutions are much more likely to seek financial aid than the public institution students.

####5.1.4.3 The median debt for students who have completed
GRAD_DEBT_MDN

@@

#6. Relationship between variables
PCTPELL vs family income 
PCTFLOAN vs family income
PCTPELL vs NPT4
PCTPELL  VS family independent
family independent   vs family income 
first generation vs family income
pell vs control
pell vs degree type
pell vs states
loan vs debt, repayment, non-repay
loan vs cost

loan vs control

Hello world

#7. Comparison between variables
PCTPELL vs PCTFLOAN

>>>>

```{r setup, include=FALSE,warning=FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE,warning=FALSE)
```

@@

library(readr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(ggvis)
library(plyr)

@@

data = read.csv("MERGED2012_13_PP.csv",header = T, na.strings = c("NULL","PrivacySuppressed"))

@@

variable = select(data, STABBR, PREDDEG, LOCALE2, MENONLY, WOMENONLY, CONTROL, ADM_RATE, SATMTMID, SATVRMID, SATWRMID, SAT_AVG,
                  NPT4_PUB,
                  NPT4_PRIV,
                  UG,
                  UG_NRA,
                  UG_UNKN,
                  UG_WHITENH,
                  UG_BLACKNH,
                  UG_API,
                  UG_AIANOLD,
                  UG_HISPOLD,
                  DEP_STAT_PCT_IND,
                  PAR_ED_PCT_1STGEN,
                  MD_FAMINC,
                  MEDIAN_HH_INC,
                  POVERTY_RATE,
                  PCTPELL,
                  PCTFLOAN,
                  GRAD_DEBT_MDN,
                  WDRAW_DEBT_MDN,
                  APPL_SCH_PCT_GE2,
                  C150_4,
                  MD_EARN_WNE_P6,
                  MD_EARN_WNE_P10,
                  RPY_3YR_RT_SUPP,
                  CDR3)

@@

```
From the output, we can tell that the highest value is 95.25% while the lowest value is 6.87%. From the mean value of 48.08%, we can tell that about for federally aided students in most colleges, around half (50.8%) of the students' parents did not complete the college. The standard deviation value of 12.2% indicated a relative high difference of this element among different colleges.

Below is the code to generate a visual representation in the histgram. 
```{R}

  ggplot(data=data_clean , aes(PAR_ED_PCT_1STGEN)) + 
    geom_histogram(breaks=seq(0.1, 1, by =0.1), 
                   col="white", 
                   aes(fill=..count..),na.rm=TRUE)+
    labs(title="Histogram for PAR_ED_PCT_1STGEN") +
    labs(x="PAR_ED_PCT_1STGEN", y="Count") +
    theme(plot.title = element_text(hjust = 0.5))
```
From the histogram, we can tell the share of first generarion students is most frequently falling into the range between 0.5 and 0.6. The value with the value more than 0.8 are rare. The colleges with the share of first generarion students value less than 0.5 are much more than the ones with the value larger than 0.5. Thus the histogram is seriously left skewed.

@@

@@

  #histogram for NPT_COM
  ggplot(data=data_clean , aes(DEP_STAT_PCT_IND)) + 
    geom_histogram(breaks=seq(0.1, 1, by =0.1), 
                   col="white", 
                   aes(fill=..count..),na.rm=TRUE)+
    labs(title="Histogram for DEP_STAT_PCT_IND") +
    labs(x="DEP_STAT_PCT_IND", y="Count") +
    theme(plot.title = element_text(hjust = 0.5))
```
From the histogram, we notice that the value between 0.7 to 0.8 are quite frequent. There was no obvious extrme value in the sample, thus the histogram is normally distributed.

@@

  ggplot(data=data_clean , aes(PCTPELL)) + 
    geom_histogram(breaks=seq(0, 1, by =0.1), 
                   col="white", 
                   aes(fill=..count..),na.rm=TRUE)+
    labs(title="Histogram for PCTPELL") +
    labs(x="PCTPELL", y="Count") +
    theme(plot.title = element_text(hjust = 0.5))
```
From the histogram, we are able to conclude that in most colleges, 40% to 50% of the undergraduate students are able to receive the Pell Grant. With little skewness, the histogram is resonalbe symmetric.

@@

  ggplot(data=data_clean , aes(PCTFLOAN)) + 
    geom_histogram(breaks=seq(0, 1, by =0.1), 
                   col="white", 
                   aes(fill=..count..),na.rm=TRUE)+
    labs(title="Histogram for PCTFLOAN") +
    labs(x="PCTFLOAN", y="Count") +
    theme(plot.title = element_text(hjust = 0.5))
```
From the histogram, it's obvious that there are two peaks in the graph. The rates between 0- 10% and the rates between 60% to 80% are similarly frequent compated to the other ranges.

@@

data_clean %>%
  group_by(variable$ctrF ) %>%
    dplsummarise(minPCTFLOAN=min(PCTFLOAN ,na.rm=TRUE),    
    maxPCTFLOAN=max(PCTFLOAN,na.rm=TRUE),medianPCTFLOAN=median(PCTFLOAN,na.rm=TRUE))  %>%
       arrange(medianPCTFLOAN)   %>%
{.}  ->PCTFLOAN_CONTROL

@@

From the result, the Private institutions have a much larger value than the public institutions. With a much higher costs burden, it makes sence that the low-income sutdents in private institutions are much more likely to seek financial aid than the public institution students.

####5.1.4.3 The median debt for students who have completed

@@

#6. Relationship between variables
#6.1 NPT4_COMBINE VS State

To analyze the relationship between average net cost for Title IV institutions and State, we grouped the data into the state level to take a further look.
```{R message=FALSE, warinig = FALSE}

data_clean %>%
  group_by(variable$STABBR) %>%
    dplyr::summarise(minNPT4_COMBINE=min(NPT4_COMBINE ,na.rm=TRUE),    
    maxNPT4_COMBINE=max(NPT4_COMBINE,na.rm=TRUE),medianNPT4_COMBINE=median(NPT4_COMBINE,na.rm=TRUE))  %>%
       arrange(desc(medianNPT4_COMBINE))   %>%
{.}  ->NPT4_STATE

#6.2 Debt VS Cost
We assume there would be a strong linear relationship between the average net price for Title IV institutions and the median debt for students who have completed. We hypothese that the students who completed education in colleges with high costs will have a highe value of median debt. We use the following code to generate the scatterplot between these two variables to verify our assumptions.

ggplot(data_clean, aes(x =NPT4_COMBINE, y = GRAD_DEBT_MDN)) + geom_point(color="red")+  geom_smooth(method = "lm", se = TRUE)+
  labs(title="Scatterplot between Cost and Debt ") +
    theme(plot.title = element_text(hjust = 0.5))

From the scatter plot, it is obvious that the NPT4_COMBINE is in positive linear relationship with GRAD_DEBT_MDN: the data show an uphill pattern from left to right; As the average net price for Title IV institutions increase, the value of median debt for students who have completed increases as well.

#6.3 Percentage of Pell Grant VS college type(CONTROL)

To compare the percentages of Pell Grant in different control type(), we grouped the Percentage of Pell Grant data into the college type level to take a further analysis.

```{R message=FALSE, warinig = FALSE}

data_clean %>%
  group_by(variable$ctrF ) %>%
    dplyr::summarise(minPCTPELL=min(PCTPELL ,na.rm=TRUE),    
    maxPCTPELL=max(PCTPELL,na.rm=TRUE),medianPCTPELL=median(PCTPELL,na.rm=TRUE))  %>%
       arrange(medianPCTPELL)   %>%
{.}  ->PCTPELL_CONTROL

ggplot(data_clean, aes(x=PCTPELL, fill=variable$ctrF)) + geom_histogram(alpha=0.9, position="identity") +
  labs(title="Overlaying histogram between PCTPELL and Control ") +
    theme(plot.title = element_text(hjust = 0.5))

```
From the overlaying histograms, the percentage of students who receive Pell Grant are quite frequent for Private For-profit institutions between 0.7 and 0.75 ; the percentage of students who receive Pell Grant are quite frequent around 0.4 for Public institutions and Private Nonprofit institutions. The histogram for private for profit is slightly left skewed due to atypical small value near 0. The histogram for public and private nonprofit institions are reasonably symmetric

#6.4  Percentage first-generation students VS Status
To compare the Percentage first-generation students with the different states, we grouped data into the state level to take a further analysis.

```{R message=FALSE, warinig = FALSE}

data_clean %>%
  group_by(variable$STABBR ) %>%
    dplyr::summarise(minPAR_ED_PCT_1STGEN=min(PAR_ED_PCT_1STGEN ,na.rm=TRUE),    
    maxPAR_ED_PCT_1STGEN=max(PAR_ED_PCT_1STGEN,na.rm=TRUE),medianPAR_ED_PCT_1STGEN=median(PAR_ED_PCT_1STGEN,na.rm=TRUE))  %>%
       arrange(medianPAR_ED_PCT_1STGEN)   %>%
{.}  ->PAR_ED_PCT_1STGEN_STATE

#6.5 family income with Pell Grant
We assume there would be a strong linear relationship between family income and the percentage of students who can receive the Pell Grant. We hypothese that colleges with small value of family income would tend to have high percentages of Pell Grant. The following codes are used to generate the scatter pkot between these two varaibles.

ggplot(data_clean, aes(x =MD_FAMINC, y = PCTPELL)) + geom_point(color="red")+  geom_smooth(method = "lm", se = TRUE)+
  labs(title="Scatterplot between MD_FAMINC and PCTPELL ") +
    theme(plot.title = element_text(hjust = 0.5))

```
From the scatter plot, it is obvious that the MD_FAMINC is in a negative linear relationship with PCTPELL: the data show a downhill pattern from left to right; As the MD_FAMINC increase, the value of PCTPELL decreases.

#7. Comparison between variables

#7.1 PCTPELL VS PCTFLOAN

library(ggplot2)
ggplot(data = data_clean) +
  geom_histogram(breaks=seq(0, 1, by =0.05),aes(x = PCTPELL, y=(..count..)), 
                       alpha=0.9, fill ="red",position="dodge") +
  geom_histogram(breaks=seq(0, 1, by =0.05),aes(x = PCTFLOAN, y=(..count..)), 
                       alpha=0.9,fill ="blue",position="dodge") +
  labs(title="Overlaying Histogram of PCTPELL and PCTFLOAN ") +
    theme(plot.title = element_text(hjust = 0.5)) + 
  scale_colour_manual("", 
                      breaks = c("PCTPELL", "PCTFLOAN"),
                      values = c("PCTPELL"="red", "PCTFLOAN"="blue"))

aid_compare -> data_clean[c(5:6)]
class(aid_compare)
